JAGS Model

### 2020-03-02 Rcodes/JAGS_overdoses.Rmd
flu over and found where I did it wrong !!!

The model defined as follows.
```{r Enter the models}
cat("model{
## define the priors
p_a ~ dbeta(alpha, beta)
mu ~ dunif(mu_a, mu_b)
sigma ~ dunif(sigma_a, sigma_b)

## the latent variables
z ~ dnorm(mu, 1/(sigma^2))
lambda <- exp(z)

for (i in 1:n) {
  ## ambulance model 
  x_a[i] ~ dbin(p_a, n_a) # each survey result for month
}
for (i in 1:n) {
  ## overdose model
  o_t[i] ~ dpois(lambda*N) # total overdoses per month
}
for (i in 1:n) {
  u_t[i] ~ dbin(p_a, o_t[i]) # ambulanced overdoses per month
}

@@

            # likelihood 
            'u_t'=u_t, 
            'x_a'=x_a,  
            'N'=N, # the population
            'n'= n, # total months
            'n_a'=n_a

            )
```
**Note: for the list object usually named 'data' or 'dat' in JAGS context, do not use arrow but use equal sign to define elements of the list.**
``` {r run the first model without error }
interations = 1000
burnin= floor(interations/2)
chains=2
# inits = list()
simple.model <- jags.model(file='basic_model.txt',
                          data=dat,
                          n.chains = chains)
simple.model
```

@@

```{r}
params= c('o_t')
samples <- coda.samples(simple.model, params, n.iter = 1000)
```

# Q1: is init necessary? and it's a characteristic of Gibbs sampling?

```{r }
summary(window(samples), start=burnin)
```

```{r include=FALSE}
plot(samples)
```

@@

```{r yield a data frame}
temp = as.matrix(samples)
colnames(temp) <- seq(1,12)
head(temp)
df_o_t <- as.data.frame(temp)
head(df_o_t)
```

```{r tidy the data frame}
df_o_t <- gather(df_o_t, key = 'month',value = 'o_t')
df_o_t$month <- factor(df_o_t$month,levels = seq(1,12))
str(df_o_t)
```

```{r}
o_t_values=data.frame('month'=seq(1,12),'o_t'=df$o_t)
o_t_values$month <- factor(o_t_values$month,levels = seq(1,12))

ggplot()+geom_point(aes(x=o_t_values$month, y=o_t_values$o_t),color='red')+geom_boxplot(aes(x=month,y=o_t),data = df_o_t)
```
### Predictive Posterior Checks

@@

ggplot()+geom_point(aes(x=u_t_values$month, y=u_t_values$u_t),color='red')+geom_boxplot(aes(x=month,y=u_t),data = df_u_t)
```

>>>>

The model defined as follows.
```{r Enter the models}
cat("model{
## define the priors
p_a ~ dbeta(alpha, beta)
mu_z ~ dunif(mu_a, mu_b)
sigma_z ~ dunif(sigma_a, sigma_b)

# unify variable names: 
# z_t <- z lmb_t <- lambda
# length unification: z_t is length 12 vector

 ## ambulance model
for (i in 1:n) {
  #Likelihood
  x_a[i] ~ dbin(p_a, n_a) # each survey result for month
}

# overdose 
for (i in 1:n) {

  ## the latent variables
  z_t[i]~ dnorm(mu_z, 1/(sigma_z^2))
  lmb_t[i] <- exp(z_t[i])

  ## overdose model
  o_t[i] ~ dpois(lmb_t[i]*N) # total overdoses per month
# Note that from pymc3 gamma was used instead of Pois dist
  u_t[i] ~ dbin(p_a, o_t[i]) # ambulanced overdoses per month
}

@@

            # likelihood 
            'u_t'=u_t, # giving data 
            'x_a'=x_a,   # giving data
            'N'=N, # the population 10000
            'n'= n, # total months 12
            'n_a'=n_a # survey size 1000

            )
```
**Note: for the list object usually named 'data' or 'dat' in JAGS context, do not use arrow but use equal sign to define elements of the list.**
``` {r run the first model without error }
chains=2
# inits = list()
simple.model <- jags.model(file='basic_model.txt',
                          data=dat,
                          n.chains = chains)
simple.model
```

@@

```{r}
params= c('o_t')
samples <- coda.samples(simple.model, params, n.iter = 1000)
# guess it's getting posterior samples ?
```

```{r }
interations = 1000
burnin= floor(interations/2)
summary(window(samples), start=burnin)
```

```{r include=FALSE}
plot(samples)
```
# q1: what is the equivalent plot that we can see we have enough iteration?

@@

```{r yield a data frame}
temp = as.matrix(samples)
colnames(temp) <- seq(1,12)
df_o_t <- as.data.frame(temp)
head(df_o_t)
```

```{r tidy the data frame}
trace_o_t <- gather(df_o_t, key = 'month',value = 'o_t')
trace_o_t$month <- factor(trace_o_t$month,levels = seq(1,12))
str(trace_o_t)
```

```{r}
# real values of the data set from pymc3 samples
o_t_values=data.frame('month'=seq(1,12),'o_t'=df$o_t)
# factorizing the months for box plot visualization
o_t_values$month <- factor(o_t_values$month,levels = seq(1,12))

ggplot()+
  # boxplot from the trace 
  geom_boxplot(aes(x=month,y=o_t), color='grey',alpha = 0.01,data = trace_o_t)+
  # real values as red dots
  geom_point(aes(x=o_t_values$month, y=o_t_values$o_t),color='red')

### Predictive Posterior Checks

@@

ggplot()+geom_boxplot(aes(x=month,y=u_t),data = df_u_t)+geom_point(aes(x=u_t_values$month, y=u_t_values$u_t),color='red')
```
